Data-Driven Modelling \& \\ Forecasting of Chaotic Systems

### 2022-01-13 Gandhi_Project_Report/draft5.tex
Changed 4.3 and H2. Conclusion writtend and added

\usepackage{enumitem}
\usepackage{multirow}
\usepackage{longtable}
\usepackage[export]{adjustbox}
% $f\arrowvert_A$ To restrict functions

@@

\author{B.L. Nortier} 
\date{January, 2022}
\date{}
\maketitle
\tableofcontents

% \begin{abstract}
%   Abstract goes here.
%   \end{abstract}

@@

The more intricate details of proofs are referred to where relevant throughout. This report is organised as follows: 
\newline In Chapter~\ref{ch2}, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE and then proceed to define the inverse-limit space and topological conjugacy of autonomous systems. 
\newline In Chapter~\ref{ch3}, we introduce the problem of forecasting dynamical systems, state the Takens delay embedding theorem and discuss various issues faced while forecasting. 
\newline In Chapter~\ref{ch4}, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this thesis/report/paper. 
\newline Finally in Chapter~\ref{ch5} we show the implementation of these forecasting methods, and conclusions are provided in Chapter~\ref{ch6}.

@@

We now turn our attention to the subject of conjugacy that describes when two dynamical systems are equivalent dynamically.

To show topological similarity or sameness between two metric or topological spaces, one needs to establish a homeomorphism between the two spaces. 
However, in the study of dynamical systems defined on two spaces, establishing a homeomorphism does not indicate that the systems are dynamically related in any way.  For instance, the maps $Tu=u^2$ and $Tu=1-|2u-1|$ defined on $[0,1]$ have totally different behaviour. 
To find systems that are dynamically similar, one rather needs to establish a kind of dynamical equivalence which we illustrate in a simple commutativity diagram below.

% Giving errors
\begin{equation}  \label{eqn_conjugacy}
%    \[ 
    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
 \everypsbox{\scriptstyle}
 \begin{psmatrix}
U & U\\%
V & V.
 %%%
%  \ncline{1,1}{1,2}^{T} 
%  \ncline{1,1}{2,1} <{\phi}
%   \ncline{2,1}{2,2}^{S}
%  \ncline{1,2}{2,2} > {\phi}
 \end{psmatrix}
% \]
\end{equation}

%Composite functions
To understand the diagram we note that if we travel ``right and then down," the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$. (right arrow downwards). Consequently the pathway amounts to finding $\phi(T(u))$. If we proceed "down and then right", the diagram instructs us to apply $\phi$ (left arrow downwards) first, and then apply $S$ (bottom arrow right). This then amounts to finding  $S(\phi(u))$. When the relationships in this diagram hold, we say $\phi(T(u))= S(\phi(u))$, and we formally denote it as $\phi \circ T=S\circ \phi$.

\begin{Definition}\rm  
[\bf {Conjugacy \& Semi-Conjugacy}]\label{Dfn_Conjugate}\rm
   Consider the dynamical systems $T:U\to{U}$, $S:V\to{V}$ and suppose the relationship $\phi \circ T=S\circ \phi$ holds. If $\phi:U\to{V}$ is a homeomorphism, then $S$ is said to be conjugate to $T$ and $\phi$ is called a conjugacy. If we relax the criterion on $\phi$ and merely require $\phi$ to be continuous and surjective where $\phi:U\to{V}$, then $\phi$ is a semi-conjugacy between $T$ and $S$ where T has domain $U$ and S has domain $V$; S is said to be semi-conjugate to T. 
\end{Definition}

\ednote{M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity. So please edit accordingly. B: Does defining explicitly the domain and codomain of $\phi$ make it clear enough? }
When $S$ is conjugate to $T$, the dynamics of the two systems are in some way ''dynamically equivalent''. Specifically, they are in one-to-one correspondence with one another. However when $S$ is semi-conjugate to $T$ with $\phi:U\to{V}$ a many-to-one mapping, the dynamics on $V$ provide merely a coarse-grained description of the dynamics on $U$ \cite{de2013elements}. When $S$ is semi-conjugate to $T$, it is also common to call $S$ a \emph{factor of $T$}, or conversely that $T$ is an \emph{extension of $S$}. In essence, an extension (e.g., \cite{de2013elements}) is a larger system capturing all of the important dynamics of its factor.

It is a very hard or nearly an impossible task to establish the existence of a conjugacy or a semi-conjugacy $\phi$ between two systems \cite{devaney2018introduction}. However, one can verify that a function $\phi$ satisfies the commutativity diagram \ref{eqn_conjugacy}.

\begin{Example}\rm
  For example $\phi(x)=\sin\big(\frac{\pi}{2}x\big)^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $U=[0,1]$.

  Indeed, $\phi:[0,1]\to[0,1]$ is a homeomorphism and by employing simple trigonometric identities, it is easily proved that $\phi$ satisfies~\ref{eqn_conjugacy}
  \[
    \begin{aligned}[t]
      (\phi\circ{T})u 
                &=\sin^{2}\Big(\frac{\pi}{2} - \frac{\pi}{2}|2u-1|\Big) = \sin^{2}(\pi{u})=\Big(2\sin(\frac{\pi}{2})\cos(\frac{\pi}{2})\Big)^{2}=4\sin^{2}(\frac{\pi}{2}u)\Big(1-\sin^{2}\frac{\pi}{2}u\Big) \\
                                    &=({S}\circ\phi)u.
    \end{aligned}
\]
And it follows that $\phi$ is a conjugacy between the two systems.
\end{Example}

In establishing that two systems are conjugate (semi-conjugate) to one another, we have also shown that one may choose to work with one system as opposed to the second and still be guaranteed to obtain information on the latter. This is especially useful while forecasting the future evolution of dynamical systems.

@@

However, the true state $u$ of a system is seldom, if ever, fully known. In almost all cases we can at most insert probes into a system to obtain partial information by means of the measurements taken. Moreover, the process of taking a measurement itself introduces 2 additional aspects which complicate the problem: \ednote{M: Maybe replace difficulty by saying that we need to consider two aspects. B: Perhaps this edit makes more sense?}
\vspace{-8mm}
\begin{enumerate}[noitemsep, label=\roman*.]
  \item A series of measurements over a specific time-interval is inherently a discretisation procedure of the underlying continuous-time dynamical system. (Hence why we restricted our attention to discrete-time systems in the preceding section)
  \item A probe will never be fully accurate, and so the act of measurement introduces a certain measure of numerical noise/inaccuracy.
\end{enumerate}

From here we construct a multidimensional observable using the method of stacking previous observations, i.e., we create delay-coordinate map defined by
$\Phi_{k,\theta}(w) := (\theta(T^{-k}w)\ldots,\theta(T^{-1}w),\theta(w))$.  
The concept of an observable is understood in the sense of an observable as originally introduced by \cite{takens1981detecting, genericObservableAeyels} to refer to a probe or measurement function inserted into the system. 
The essence of Takens' theorem relates to the fact that when $k$ is sufficiently large, we can define a dynamical system on the space $\mathbb{R}^{k+1}$ whose states are $\Phi_{k,\theta}(w)$, $\Phi_{k,\theta}(Tw)$, $\Phi_{k,\theta}(T^2w)$, etc. and this dynamical system is topologically conjugate to the unknown underlying system $(W,T)$. We recall the Takens delay embedding theorem next.

@@

\section{Takens Embedding Theorem}\label{sect_Takens}

We define first the concepts of homeomorphism and embedding:
\begin{Definition}\rm
  [\bf {Homeomorphism}]\label{Dfn_homeo}\rm
  A homeomorphism is a function $f:Z\rightarrow Y$ between two topological spaces $Z$ and $Y$ that is continuous, bijective and has a continuous inverse. 
\end{Definition}

\begin{Definition}
  [\bf {Embedding}]\label{Dfn_embed}\rm
  Consider a homeomorphism $f:Z\rightarrow Y$ for $Y\subset X$. $Z$ is said to be embedded in $X$ by $f$.
\end{Definition}

Takens Theorem states a result establishing a relationship between the observed and underlying dynamical systems by showing that the concatenation of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.  We formulate the theorem from \cite{takens1981detecting}.

\begin{Theorem} 
	[\bf Takens Embedding Theorem (adopted from \cite{takens1981detecting}] \label{Thm_Takens}
         Let $W$ be a compact manifold of dimension $m$, and $d\ge m$ so that $2d$ is an integer. It is a 
            generic property for the pair $(T, \theta)$,  where $T:W \to W$ is
            a smooth diffeomorphism, and $\theta:W \to \mathbb{R}$ a smooth function, the map $\Phi_{2d,\theta}:W \to \mathbb{R}^{2d+1}$ defined on $W$ by 
            $\Phi_{2d,\theta}(w) := (\theta(T^{-2d}w)\ldots,\theta(T^{-1}w),\theta(w))$
            is a diffeomorphic embedding; by `smooth' we mean at least $C^2$. Consequently, there exists a map $F_\theta: \Phi_{2d,\theta}(W) \to \Phi_{2d,\theta}(W)$ defined by $$F_\theta: (\theta(T^{-2d}w),\ldots,\theta(T^{-1}w),\theta(w)) \mapsto 
            (\theta(T^{-2d+1}w),\ldots,\theta(w),\theta(Tw))$$
           so that $(W,T)$ is topologically conjugate to 
            $(\Phi_{2d,\theta}(W), F_\theta)$.    
\end{Theorem}

By generic we mean a residual set on a certain topology on appropriate spaces of functions (that we do not describe here) \textbf{(Cite)}.  %\ednote{How vague may one be in general by leaving things out?} 
By $C^2$, we make reference to a twice-differentiable function  with a continuous second derivative. In our scenario (and this remains important throughout) the input space $U$ is considered the attractor.~\label{attractor_U}

Below we reproduce Figure 1 from \cite{Supp}.

%Some explanation is beneficial. Takens establishes a delay-coordinate map $\Ftheta$ defined in \ref{eqn_takens} for $T$ the flow (\ref{defn_flow}) defined by the update equation  $Tu_n=u_{n+1}$ and $theta$ our measurement function. When confining a dynamical system to the manifold U (in which the underlying system is contained), Takens showed that if certain smoothness conditions are satisfied on $T$ and $\theta$, then the delay-coordinate map $\Ftheta$ embeds $U$ in the reconstruction space $R^{2d+1}$ \textbf{we haven't used the term reconstruction space yet, so this is falling out of the blue}, for almost every choice of measurement function $\theta$, the observable. By almost every choice here, we mean that we exclude some observations which do not give any information. To see a more precise discussion, see \textbf{xxx} for we do not concern ourselves with a further elaboration on this report.

%Alternatively we may also make the statement: Takens' Embedding Theorem guarantees that almost all dynamical systems can be reconstructed from just one noiseless observation sequence \textbf{this isn't obvious from the theorem's formulation, is it?} i.e. for a great number of possible observation functions $\theta$, $\Ftheta$ preserves the topology of U. \textbf{'preserving the topology' is not something we've touched on before. Perhaps I should add to previous discussion so logic flows seamlessly here}.

\begin{figure}[ht]
  \includegraphics[scale=0.25]{_takensmap.eps}
  \centering
  \captionof{figure}{Schematic of Takens Embedding Theorem's usage and consequence of how iterates \ldots }
  \label{fig:takensmap}
\end{figure}
\ednote{M: Here, and throughout if you reproduce a figure, it must be stated where it was reproduced from. B: Is this satisfactory? Also, should $\tilde{F}$ be left out for now?}
\ednote{B: This is only a screenshot used as a temporary placeholder. If sufficient, I shall convert it to .eps. If not, may I perhaps have the original picture?}

If we can learn the map $F_\theta$, from sufficient number of data points of $\{\phi_{2d,\theta}(w_n)\}$, then we also know how to forecast how $\theta(w_n)$.

%Once again we pause and consider the graph to consolidate our understanding of the conjugacy (\textbf{are we being too simplistic?}). If we are in state $u$ and then evolve towards state $Tu$ before embedding into $\mathbb{R}^{2m-1}$, then this is equivalent to first embedding into $\mathbb{R}^{2m-1}$ and then evolving through the map $\Ftheta$, i.e. $\Phi_{2d,\theta}\circ{F}_{\theta} = T\circ\Phi_{2d,\theta}$. The map $F_theta$ is a homeomorphism as well.

%Thus information about U can be retained in the time series' (or observation's) output. By preserving the topology on the manifold $U$ in the reconstruction space $X$ (\textbf{not discussed before}), we are also guaranteed the preservation of topological invariants of the manifold, of which dimensionality is one such invariant. (We note here that dimensionality will again be considered later in this project).

\section{Practical Limitations to Takens' Theorem}

It is a fact that Takens' Embedding Theorem is both a powerful result and that it provides compelling reason to believe that one could conceivably reconstruct accurately a system conjugate to the underlying system. Nonetheless, it does present some serious practical limitations:
\vspace{-5mm}
\begin{enumerate}
\item Even supposing that we could indeed find $F_\theta$ , our \emph{approximation} of  $F_\theta$ is a map from a larger set $\mathbb{R}^{2d+1}$ containing the embedded attractor. There are, however, no theoretical guarantees that $F_\theta$ will retain  $V=\Phi_{2d,\theta}(W)$  as an attractor although $W$ could itself be an attractor.
\item Takens theorem is stated only for noiseless observations. Due to noise $\epsilon_n$ the delay vector $\Phi_{2d,\theta}(w_n) + \epsilon_n$  may lie outside $V$.Furthermore, due to the chaotic nature of the underlying system (i.e. the fact that it has SIDC), the evolution of $\Phi_{2d,\theta}(w_n) + \epsilon_n$ under the map $F_\theta$ could move out of $V$ completely. This problem can be overcome  by using a driven dynamical system with some properties, and we discuss this in Chapter\ref{ch3}. 
\end{enumerate}

% Another limitation owes itself to the requirement that T be a surjective map. This is not always the case in practical examples and we may consider the case of a damped double pendulum \ednote{M: Please remove the paragraph as well. We do not overcome this problem. You can add this in the next Chapter where you need T to be surjective or else you could include this at the begining of this chapter where you explain the forecasting problem.}

The Takens embedding theorem does not guarantee global dissipativity.
\ednote{B:Links with~\ref{subs_LearnGamma} in discussion of dissipativity of $\Gamma$.  Sir, I'm not sure how to write this here - I think it a beneficial point to make so as to set the stage for the later discussion in ~\ref{subs_LearnGamma}, but I am unsure there as well.}

\ednote{Include note on learning failure because of sensitivity as in graph commented out?}
% \begin{figure}[ht]
%   \includegraphics[scale=0.3]{_learningfailure.eps}
%   \centering
%   \caption{Illustration of a system having failed to learn the map}
%   \label{fig:supp_learning_failure}
% \end{figure}

Though not a uniquely related to the Takens embedding theorem, we do opt to make mention of this requirement for our work. 
The methods described here are applicable for data originating from a surjective map. 
This follows from the fact that when $T$ maps the space $U$ into a proper subset $B$ of $U$, we me conclude that $T$ is not surjective as there exists some point in $U-A$ which does not have a preimage.
An example of this would be the case when a system displays energy-loss. Consider, for instance, consider the oscillations dying down for a damped pendulum. (A more in-depth account will be given in \ref{ch5})

This requirement is not overly restrictive, as a great number of chaotic systems do have a surjective map. (Cite) \ednote{I rememeber you saying this in one of our sessions, but don't have a source for it.}

\chapter{Driven Dynamical Systems to Forecast Problems}\label{ch4}

In this chapter, we discuss results pertaining to the mapping of temporal data obtained from a discrete-dynamical system onto a different space through the notion of a driven dynamical system. 
We also consider the conditions a driven system should possess to avoid adding distortion to its state-space representation. We then describe the \emph{single-delay dynamics} (SDD) of the system and and finally recall conditions on the driven system so that the SDD are conjugate (or at least semi-conjugate) to the underlying system. 
The SDD can then be used to forecast and reconstruct the underlying system. 
% \ednote{Is this attractor assumed to exist? M:  I replaced  it with the system. Attractor is also a commonly used term since we observe data from an attractor in almost all cases.}

\section{Nonautonomous and Driven Dynamical Systems}

\begin{Definition}
  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP}))] \label{Dfn_NDS}\rm
  A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps $\{g_n\}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous. Each $g_n$ arises as a consequence of an input $u_n$ from the input space $U$, a topological space. 
\end{Definition}

We immediately recall the figure first defined in \eqref{eqn_conjugacy}  and remind ourselves of our ultimate goal to learn a system that is conjugate to an underlying or unknown dynamical system $(U,T)$ by making use of the measurements obtained from  the unknown system.

%#####################################
\begin{center}
\psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
\everypsbox{\scriptstyle}
\begin{psmatrix}
U & U\\%
V & V.
%%%
%  \ncline{1,1}{1,2}^{T} 
%  \ncline{1,1}{2,1} <{\phi}
%   \ncline{2,1}{2,2}^{S}
%  \ncline{1,2}{2,2} > {\phi}
\end{psmatrix}
%#####################################
\end{center}

To find a function $\phi$, the need now arises to consider a so-called driven dynamical system, a special case of a nonautonomous dynamical system, where two spaces are considered: input is taken from an input space $U$ and the state of a space $X$ is updated. (Previously, only the state was considered.)
We do this to mimic the true conditions. By accounting for both external and internal factors which influence the  system-state at a specific time-step, our model resembles nature much more realistically.  \ednote{M : I understand this sentence, but am not sure if a general audience would. So please elaborate or omit. B: Is this satisfactory?}

\begin{Definition}
  [\bf Driven, Compactly Driven Dynamical System] \label{Dfn_DDS} \rm
A driven dynamical system comprises two topological metric spaces $U$, $X$ and a continuous function  $g:U\times{X}\to{X}$ where $g(u_n, x_n)=x_{n+1}$.
If the input space $U$ is compact, we refer to the system as compactly driven. 
\end{Definition}

The dynamics on X are generated by the update equation $x_{n+1}=g(u_n, x_n)$ where $n\in\mathbb{Z}$, input $u_n$ from $U$ and state $x_n$ belonging to $X$, where $X$ is compact.
Abbreviated, we shall refer only to the \emph{driven system $g$}, with all other entities being understood implicitly.

Notably, a nonautonomous dynamical system may be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps $\{g(u_n, \cdot)\}_{n\in\mathbb{Z}}$ contained in $X$.
Physically, one may think of this bi-infinite sequence as referring to a system that has been running for a long period at the time of the first measurement taken from the system (alternatively, the first time a probe is inserted into the system to take an observation)

\begin{Definition}
  [\bf Entire Solution] \label{Dfn_Soln} \rm
  A sequence $\{x_n\}_{n\in\mathbb{Z}}\subset X$ is called an entire solution (or simply a solution) of the driven system  $g$ with input $\overline{u}$ when it satisfies 
  \[g(u_{n-1}, x_{n-1})=x_n\] for all $n\in\mathbb{Z}$
\end{Definition}

It is important to emphasise that a sequence $\{x_n\}$ satisfying the update equation above can only be a solution if $x_n\in{U}$ holds for all $n\in\mathbb{Z}$ and just for $n>0$. Consider the example below.

\begin{Example} \rm \label{ex_halfux}
  The only solution  $\{x_n\}_{n\in\mathbb{Z}}$ to the driven system  $g(u,x)=\frac{ux}{2}$, where $X=[0,1]$, $U=[0,1]$,  is the zero solution $x_n\equiv0$.
  To see this, consider any $x_n=a\in[0,1]$ where $a\neq{0}$.  Let $\overline{u}\in{U}$ be an non-zero constant sequence , say $u_n=0.5$. 
  The driven system may be rewritten as $x_{n-1}=\frac{2x_n}{u_{n-1}}=4x_n$ and the  iterates of $x_n$ in backward time will increase by a factor of 4 at each timestep. 
  Thus for some $m\leq{n}$,  $1<x_m$ i.e. $x_m\notin{X}$. So ${\{x_n\}}_{n\in\mathbb{Z}}$ is not a solution and it follows that the only possible solution is the zero solution.
\end{Example}
% \ednote{Nicely written!}

A system may also have multiple solutions as is evidenced in the example below.

\begin{Example}\label{Ex_exp} \rm 
 Consider the driven system $g(u_n,x_n)=x^n$ for $X=[0,1]$, $U=\mathbb{R}$. The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $lim_{n\to\infty}x_n=0$, $lim_{n\to{-}\infty}x_n=1$. 
 The proof is deferred to immediately after the next paragraph (see~\ref{rem_proofEx}).
\end{Example}

As the solutions to a driven system are often considered to be an important entity \cite{M: edited the sentence here}, we next identify a subspace $X_U$ of $X$ that contains all possible solutions. To realize such a subspace of a driven system $g$, the concept of a reachable set is defined.

\begin{Definition}
  [\bf Reachable Set]\label{Dfn_ReachableSet}\rm
The reachable set of a driven system $g$ is exactly the union of all the elements of all the (entire) solutions, i.e., 
\[X_U :=\Big \{x \in X:  x = x_k \mbox{ where  $\{x_n\}$  is a solution for some  $\bar{u}$} \Big \}.\]
The set of all reachable states at a specific time $n$ for input $\overline{u}$ is denoted by $X_n(\overline{u})$
\end{Definition}

The reachable set itself can defined independently from the fact that $g$ does or does not have the ESP/USP. We note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$. \textbf{(Cite)}.

$X_n$ is exactly the set of points $x$ through which some entire solution $\Psi$ passes at time $n$.

This will lead to a result established later, but which is worth taking note of now: \textit{$g$ being a topological contraction is equivalent to the existence of a unique entire solution.}

% \begin{Definition}
%   [\bf Topological Contraction]\label{Dfn_TopContr}\rm
%   A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subset{U}$, $X_n(\overline{u})$ is a singleton subset of $X$.  
% \end{Definition}

\begin{Remark}
  [\bf Proof of Example~\ref{Ex_exp}] \label{rem_proofEx} \rm
  To see this, we show that for every input $\overline{u}=\{u_n\}_{n\in\mathbb{Z}}$, $g(u_n,\cdot)$ is a contraction map on $X$ for every $k\in\mathbb{Z}$. 
  Indeed $g(u_n,x_n)=x^n$ depends only on the state at time $n$ and so for each input $\overline{u}$, $X_n(\overline{u})=\{x^n\}$ is a singleton subset.
  We then easily conclude that $\{x_n\}\equiv0$ is the only solution. 
\end{Remark} \ednote{B: Is this sufficient?}

Thus far it has been demonstrated that a system may have one or more solutions; one may ask if a driven system always has a solution and, if so, whether it satisfies certain properties such as uniqueness. 
Should the driven system be compact, existence follows immediately as shown in the following result.

\begin{Theorem}\label{Thm_CompactExistence}
 Let $g$ be a driven system.  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
\end{Theorem}
\begin{proof}
  May be found in \cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
\end{proof}

One may easily construct many systems with trivial solution-sets, such as $g(u,x)=x$ which has only the constant solution $x$ and so for $U=[-1,1]$, the system would have no solution if $|x|>1$. To refine the scenario, we consider only systems with unique solutions.

\section{Unique Solution Property}

\begin{Definition}
  [\bf Unique Solution Property] \label{Dfn_usp}\rm
  A driven system $g$ is said to have the Unique Solution Property (USP) if for each input $\overline{u}$ there exists exactly one solution. 
  Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$, with $\Psi({\overline{u}})$ denoting the unique solution.
\end{Definition}

One of the first result obtained after defining the USP is the fact that every solution will attract different initial conditions towards the component parts of the solution. %, i.e. that the solution $\Psi$ is a non-autonomous uniform attractor. \ednote{Define Uniform Attractor before stating theorem or else omit the statement.} 
If $g$ has the USP, then any solution to $g$ is also a uniform attractor in nonautonomous dynamical systems literature \cite{Manju_Nonlinearity}. The discussion on nonautonomous attractors is beyond the scope of this project and we refer the reader to \cite{Manju_ESP, esann2012ids} for further reading.

% \ednote{M: You can omit the theorem, and just say it is a uniform attractor in the nonautonomous dynamical systems literature \cite{Manju_Nonlinearity}, and say the discussion on nonautonomous attractors is beyond the scope of the thesis.}
% Paraphrased.
Additionally, $g$ having the USP is in our context also equivalent to $g$ being a topological contraction \cite{manjunath2021universal}, or alternatively also that $g$ exhibits the Echo State Property (ESP).

In the majority of Reservoir Computing(RC) literature (a machine learning algorithm approach)\ednote{How should I explain/qualify RC better here? It's mentioned without really explaining.}, 
the ESP is used to describe the concept of fading memory or memory-loss  \cite{jaeger2004harnessing, lu2018attractor}. 
If $g$ possesses the ESP (equivalent in our context to the USP), the we are guaranteed that the whole of an input's left-infinite history will exactly determine the current system state; i.e. there is only ever the possibility of a single reachable state at a given instance of time \cite{jaeger2001echo}

Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical benefit since a relationship between the network's response and the temporo-statistical properties of the input are not directly described. A more detailed presentation may be found in~\cite{jaeger2001echo}.

We do however consider it important to mention that the ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks(RNN) within the field of RC \cite{Manju_ESP}.

\section{Next section?}
\ednote{I feel a break is necessary here, but would hear from you what your thoughts are.}

Having already defined the reachable set $X_U$ as the set of all elements of all entire solutions, we pause briefly in order to fix additional notation.
Defining $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ as the left-infinite subsequence of an input up until time $n$, $\overleftarrow{U}$ is then the notation for all these left-infinite sequences in $U$. 
Moreover, $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ will symbolise the input up to time $n$ with $v \in U$ being the specific input value at time $n$. 
The introduction of a new input at time $n$ can be described by the mapping $\sigma_v:   \cev{u}^{n} \mapsto \cev{u}^{n}v$. 
The right-shift map, $r\overleftarrow{U}\to\overleftarrow{U}$, of an input sequence is defined $r: (\cdots, u_{-2},u_{-1}) \mapsto(\cdots, u_{-3},u_{-2})$

The goal now assumes the form of the question on our ability to establish a semi-conjugacy as presented below for the driven system $g$. 
%\ednote{M: I have changed it to a semi-conjugacy since conjugacy amounts to asking a infinite-dimensional space $\overleftarrow{U}$ to be embedded in $X_U$ which is not possible.}

\begin{equation}  \label{Scomm_h}
  %    \[ 
      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
   \everypsbox{\scriptstyle}
   \begin{psmatrix}
   \overleftarrow{U} & \overleftarrow{U}\\%
   X_U & X_U.
   %%%
  %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
  %  \ncline{1,2}{2,2} > {h}
  %  \ncline{2,1}{2,2}^{g(v,\cdot)}
   \end{psmatrix}
  % \]
  \end{equation}

We proceed to consider a specific subclass of conjugacies.

  \begin{Definition}
    [\bf Universal Semi-Conjugacy]\label{Def_UnivSemiConj} \rm
    Given a driven system $g$, we  call a continuous and surjective map $h : \overleftarrow{U} \to X_U$ a universal semi-conjugacy if  diagram \ref{Scomm_h} commutes for all $v \in U$.
  \end{Definition}

If the universal semi-conjugacy $h$ exists (i.e. the diagram in \ref{Scomm_h} commutes) then the solution $\Psi$ will be at least a coarse-grained representation of the input $u$. \ednote{B: Expand/Cite}

But does such a function $h$ for the above driven system $g$ above exist? Whenever $g$ has the USP and $\Psi(u)=\{x_n\}_{n\in\mathbb{Z}}$ it follows that $h$, defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$, will satisfy the semi-conjugacy in the graph above \ref{Scomm_h}.
Regrettably, such a continuous mapping $h$ is not guaranteed to exist \cite{M: Have edited here} when $g$ does not have the USP \cite[Lemma 5]{Manju_Nonlinearity}.
Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example \ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.

Re-sketching the commutativity diagram \ref{Scomm_h} above by fixing the input $v$ in $g$ and replacing $X_U$ by its left-infinite sequence space $\overleftarrow{U}$, we obtain the diagram below. In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping}.

\begin{Definition}
  [\bf Causal Mapping]\label{Def_CausMap}
  A continuous, surjective map $H:\overleftarrow{U}\to\overleftarrow{X}_U$ such that \[H\circ\tilde{g}_v=\sigma\circ{H}\] where $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$ is called a causal mapping.
\end{Definition}

>>>>

\usepackage{enumitem}
\usepackage{multirow}
\usepackage{longtable}
\usepackage[export]{adjustbox}

@@

\author{B.L. Nortier} 
\date{January, 2022}
\date{}
\maketitle
% \tableofcontents

@@

The more intricate details of proofs are referred to where relevant throughout. This report is organised as follows: 
\newline In Chapter~\ref{ch2}, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE and then proceed to define the inverse limit space and topological conjugacy of autonomous systems. 
\newline In Chapter~\ref{ch3}, we introduce the problem of forecasting dynamical systems, state the Takens delay embedding theorem and discuss various issues faced while forecasting. 
\newline In Chapter~\ref{ch4}, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this thesis/report/paper. 
\newline Finally in Chapter~\ref{ch5} we show the implementation of these forecasting methods, and conclusions are provided in Chapter~\ref{ch6}.

@@

@@

However, the true state $u$ of a system is seldom, if ever, fully known. In almost all cases we can at most insert probes into a system to obtain partial information by means of the measurements taken. Moreover, the process of taking a measurement itself introduces 2 additional aspects which complicate the problem: 
\vspace{-8mm}
\begin{enumerate}[noitemsep, label=\roman*.]
  \item A series of measurements over a specific time-interval is inherently a discretisation procedure of the underlying continuous-time dynamical system. (Hence why we restricted our attention to discrete-time systems in the preceding section)
  \item A probe will never be fully accurate, and so the act of measurement introduces a certain measure of numerical noise/inaccuracy.
\end{enumerate}

From here we construct a multidimensional observable using the method of stacking previous observations, i.e., we create delay-coordinate map defined by
$\Phi_{k,\theta}(w) := (\theta(T^{-k}w)\ldots,\theta(T^{-1}w),\theta(w))$.  The essence of Takens theorem relates to the fact that when $k$ is sufficiently large, we can define a dynamical system on the space $\mathbb{R}^{k+1}$ whose states are $\Phi_{k,\theta}(w)$, $\Phi_{k,\theta}(Tw)$, $\Phi_{k,\theta}(T^2w)$, etc. and this dynamical system is topologically conjugate to the unknown underlying system $(W,T)$. We recall the Takens delay embedding theorem next.

@@

\section{Takens Embedding Theorem}

% \section{Practical Issues of Takens theorem}

% \chapter{Driven Dynamical Systems to Forecast Problems}\label{ch4}

% In this chapter, we discuss results pertaining to the mapping of temporal data obtained from a discrete-dynamical system onto a different space through the notion of a driven dynamical system. 
% We also consider the conditions a driven system should possess to avoid adding distortion to its state-space representation. We then describe the \emph{single-delay dynamics} (SDD) of the system and and finally recall conditions on the driven system so that the SDD are conjugate (or at least semi-conjugate) to the underlying system. 
% The SDD can then be used to forecast and reconstruct the underlying system. 
% % \ednote{Is this attractor assumed to exist? M:  I replaced  it with the system. Attractor is also a commonly used term since we observe data from an attractor in almost all cases.}

% \section{Nonautonomous and Driven Dynamical Systems}

% \section {Unique Solution Property}

% % \section{Next section?}

% \section{Choosing the driven system $g$}

% \section{The next step in Dynamics}

% % \section*{Summarising the discussing thus far:}

% \section{A discussion of $G_T$ }

% \section{Advantages of learning $\Gamma$} \label{sect_LearnGamma}

% \chapter{Implementation} \label{ch5}

% In this chapter we turn to the implementation of the function $G_T$ by discussing a method to learn the function $\Gamma$. We then indicate where this implementation differs from previous work done by (\cite{manjunath2021universal}) and empirically show that the SDD of $G_T$ are less functionally complex than $\Phi_{k, \theta}$ or $T$ by considering the Pearson correlation coefficient. 
% Finally, we present and discuss some numerical results for a focus system of this project: the double pendulum. We also consider other chaotic attractors.

% \section{Implementing $G_T$}

% \section{Changes in this project to previous work of~\cite{manjunath2021universal}}

% \section{Numerical simulations}
% % In this section we showcase numerical results pertaining to 3 attractors that were simulated using the methodology described in the previous chapters. We first consider a simple physical system.
% \subsection{Double Pendulum}
% \subsection{Alternative Attractors}
% \subsubsection{Thomas' Cyclically Symmetric Strange Attractor}
% \subsubsection{Fractal Dream Attractor}
% \section{Generalised Pearson Correlation} %\ednote{B: Alter heading.}
